Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-19 article.tex
update experimental sction

As the first plot in Figure~\ref{fig:filtered-signal} illustrates, the raw
fluorescence signal is very noisy due to light scattering artifacts that
ordinarily affects the quality of the recording~\citep{lichtman2011big}.
Accordingly, the first step of our pipeline is to smoothen the signal, using
one the following low-pass filter for filtering out high frequency noise :
\begin{align}
% symmetrical median filter
f_1(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i \label{eq:symetric-median}, \\
% asymmetrical weighted median
f_2(x^t_i) &= 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_{t}^i,
\label{eq:weighted-asymetric-median}
\end{align}
where \textcolor{red}{$f_?$} effect is shown in the second plot in the figure.

@@

% Datasets
To assess our method, we focus on the \textit{normal-1} dataset
provided by the organizers of the Connectomics challenge. Experiments
were also carried out on  \textit{normal-2, normal-3, normal-4} and show
similar trends.
Those datasets have $p=1000$ time series of length $T=179500$ of image
calcium from simulated neuronal networks \citep{stetter2012model}. Those
networks have approximately 15000 edges.

@@

Two reasons might explain such overwhelming difference: (i) among
those methods only the partial correlation makes the distinction between
direct and indirect connections, (ii) the simulated network from the
\citep{stetter2012model} generator could be reduced to a linear model after
pre-processing. Those results is obviously generalized over the other
datasets see Appendix \textcolor{red}{REF}.

As it can also be seen on Figure \ref{fig:curves}, the GENIE3 performance
increases with deeper trees in addition to be computationally feasible. The
reason is twofold. First, the deeper the tree, the larger is the conditioning.
As it was observed in the comparison between correlation and partial
correlation, it screens indirect links in the inferred network. Second, the
number of small trees need to be much higher than deep trees in order to
observe - and evaluate - all pairs.  However, as suggested by
\cite{louppe2014understanding}, limiting the maximal depth of trees may help
to reduce the misestimation bias in variable importance computing, i.e.
inferring the strength of a link. Therefore a trade-off of the maximal depth may give the best of both worlds.

@@

\begin{figure}[bth]
\centering
\subfigure[]{\includegraphics[width=0.45\textwidth]{images/curve_roc} \label{fig:roc_curve}}
\subfigure[]{\includegraphics[width=0.45\textwidth]{images/curve_pr} \label{fig:pr_curve}}
\caption{Method precision-recall and ROC curves. \textcolor{red}{GENI3?}
         $\tau=0.11$}
\label{fig:curves}
\end{figure}

@@

\subsection*{Signal processing highly improves performance}

\textcolor{red}{The whole sub-section should be moved to the appendix and
                replace by one or two sentences.}

@@

>>>>

As the first plot in Figure~\ref{fig:filtered-signal} illustrates, the raw
fluorescence signal is very noisy due to light scattering artifacts that
ordinarily affects the quality of the recording~\citep{lichtman2011big}.
Accordingly, the first step of our pipeline is to smoothen the signal, using
one the following low-pass filter for filtering out high frequency noise :
\begin{align}
% symmetrical median filter
f_1(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i \label{eq:symetric-median}, \\
% asymmetrical weighted median
f_2(x^t_i) &= 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_{t}^i,
\label{eq:weighted-asymetric-median}
\end{align}
where $f_1$x effect is shown in the second plot in the figure.

@@

% Datasets
To assess our method, we focus on the \textit{normal-2} dataset
provided by the organizers of the Connectomics challenge. Experiments
on \textit{normal-1, normal-3, normal-4} show similar trends to
\textit{normal-2} and are presented in the Appendix \textcolor{red}{REF}.
Those datasets have $p=1000$ time series of length $T=179500$ of image
calcium from simulated neuronal networks \citep{stetter2012model}. Those
networks have approximately 15000 edges.

@@

\begin{figure}[bth]
\centering
\subfigure[]{\includegraphics[width=0.45\textwidth]{images/curve_roc} \label{fig:roc_curve}}
\subfigure[]{\includegraphics[width=0.45\textwidth]{images/curve_pr} \label{fig:pr_curve}}
\caption{The weighted average of partial correlation matrix, winning approach
         of the challenge, yields better than AUROC and AUPRC than all
         other state-of-the-art methods. (parameters: $\tau=0.11$, partial
         correlation used filer $f_1$).}
\label{fig:curves}
\end{figure}

Two reasons might explain such differences: (i) among
those methods only the partial correlation makes the distinction between
direct and indirect connections, (ii) the simulated network from the
\citep{stetter2012model} generator could be reduced to a linear model after
pre-processing. Those results is obviously generalized over the other
datasets see Appendix \textcolor{red}{REF}.

\textcolor{red}{AJ: GENI3 is not on the figure}
As it can also be seen on Figure \ref{fig:curves}, the GENIE3 performance
increases with deeper trees in addition to be computationally feasible. The
reason is twofold. First, the deeper the tree, the larger is the conditioning.
As it was observed in the comparison between correlation and partial
correlation, it screens indirect links in the inferred network. Second, the
number of small trees need to be much higher than deep trees in order to
observe - and evaluate - all pairs.  However, as suggested by
\cite{louppe2014understanding}, limiting the maximal depth of trees may help
to reduce the misestimation bias in variable importance computing, i.e.
inferring the strength of a link. Therefore a trade-off of the maximal depth
may give the best of both worlds.

@@

@@

@@

\section{Signal processing highly improves performance}
